### 2024-12-03 Bias-in-arrests-toronto/paper/paper.qmd
Initial Commit

---
title: "Analyzing Arrest Patterns and Racial Representation in Toronto"
author: "Akshat Aneja"
abstract: This paper analyzes arrest data from the Toronto Police Service for the years 2020 and 2021, focusing on the distribution of arrests by ethnic identity, age, and booking status to assess whether systemic bias exists in policing practices. The analysis reveals that, while arrests appear to be driven by objective factors, disparities related to socio-economic status may still exist. These findings provide important insights that can help improve transparency and accountability in policing, contributing to more equitable law enforcement practices.
date: today
date-format: long
number-sections: true
format: pdf
toc: true
thanks: "Code and data supporting this analysis is available at: https://github.com/Akshat211202/Analysis-of-bias-in-arrests/"
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(opendatatoronto)
library(dplyr)
library(tinytex)
library(ggplot2)
library(knitr)
library(kableExtra)
library(modelsummary)
library(bayesplot)
library(rstanarm)
```

```{r}
#| include: false
#| warning: false
#| message: false

cleaned_data <- read.csv(here::here("data/02-analysis_data/analysis_data.csv"))

# import model
model <- readRDS(here::here("models/racial_model.rds"))
```

# Introduction

This study examines disparities in booking probabilities across ethnic groups in Toronto during 2020 and 2021. While discussions about racial bias in policing often focus on arrest rates or use of force, booking decisions represent another critical stage in law enforcement processes that can reflect systemic bias.

Booking, the formal process of charging an individual after arrest, determines whether someone enters the criminal justice system. Examining booking probabilities provides insight into whether disparities exist in how different racial groups are treated after an arrest. This analysis focuses on identifying patterns in booking decisions for various ethnic groups and whether these patterns changed over the two years.

Previous studies on policing in Toronto have reported that individuals from Black, Indigenous, and other minority groups are overrepresented in police actions relative to their population size. However, most of these studies focus on arrest rates or the use of force rather than the probability of being booked. This study seeks to address that gap by analyzing the relationship between ethnicity and booking status.

The study uses arrest and booking data from 2020 and 2021, applying Bayesian logistic regression to model the likelihood of being booked after arrest. The analysis includes ethnicity and year as predictors. The goal is to determine whether booking practices show evidence of systemic bias or reflect consistent patterns across groups.

The overview of the data set, its preparation and summary statistics are in [@sec-data]. The description of the Bayesian logistic regression model used for analysis, including justification for its use is in [@sec-model]. Presentation and interpretation of the model outputs, highlighting key patterns and trends are in [@sec-results]. Analysis and summary of findings in the context of existing research, limitations of the study and ecommendations for future research are in [@sec-discussion].

The findings aim to provide objective evidence that can inform discussions about equity in policing. Identifying disparities in booking probabilities can help highlight areas where procedural changes may be necessary to ensure fair treatment for all groups. This study also explores how booking practices have evolved between 2020 and 2021 to identify potential progress or emerging trends.

# Data {#sec-data}

This data is sourced from the Toronto Open Data Portal `opendatatoronto` [@rOpenDataToronto]. The following data is used for an in-depth analysis of arrests and strip searches which involve various ethnic and identity-based groups in the city of Toronto covers the period from 2020 to 2021. The information is gathered by the Toronto Police Service (TPS) [@torontoPolice] and is collected under the authority of the Police Services Act.Instrumentation for this process involved the direct input of data by law enforcement officers during or after arrest events, ensuring the capture of relevant variables tied to each arrest.

The raw dataset comprised 32,000 arrest records from 2020 and 2021, with 26 variables available for analysis which reflects real-time operational data collected through TPS's digital reporting systems, designed to improve transparency and accountability in policing practices. For this study, we focused on specific variables, including Arrest Year, Perceived Race, Age Group, Youth at Arrest, Booked, and the various Search Reasons (Cause Injury, Assist Escape, Possess Weapons, and Possess Evidence). To simplify the analysis, we consolidated these search-related columns into a single "Search Reason" column using the `any()` function [@rDplyr].

The dataset represents real-world phenomena, such as arrests and bookings, as entries recorded by law enforcement officers during or after these events in Toronto for 2020 and 2021. Key variables include perceived_race (officer-assessed racial identity), year (2020 or 2021), and booked (whether the individual was formally charged, a binary variable). Arrest counts are used to calculate booking probabilities (bookings/arrests). While the dataset enables analysis of booking trends by ethnicity and year, some limitations exist. Perceived race is subjective and may introduce bias, contextual factors such as offense type or socio-economic status are absent, and the binary nature of the booking variable simplifies a complex decision-making process. Despite these constraints, the dataset provides a structured basis for investigating equity in booking practices.

The data underwent cleaning and analysis using the R programming language [@citeR]. Cleaning was performed with the `tidyverse` package [@rTidyverse]. Subsequently, analysis was conducted utilizing the `dplyr` package [@rDplyr], visualization was done using the `ggplot2` package [@rGgplot2], depicting the number of arrests and strip searches across various ethnic and identity-based groups.

## Measurement

This study aims to estimate the extent of racial bias in arrest patterns across different ethnic groups in Toronto for the years 2020 and 2021. The key variables of interest include Perceived Race, Age Group, Youth at Arrest, and Booking Status. The primary estimand is the relationship between perceived race and arrest outcomes, specifically whether individuals from marginalized communities are disproportionately arrested or booked relative to their representation in the population.

The analysis is conducted using the arrest data as the primary estimator. Arrest records are grouped by ethnic identity, age, and booking status to calculate the distribution of arrests across these categories. The study also considers broader socio-economic factors such as poverty and unemployment, which may affect arrest rates, as potential confounders. By quantifying these relationships through statistical analysis and data visualization, the study seeks to determine whether arrest practices reflect systemic bias or are driven by objective factors.

## Total number of arrests by age group for the year 2020 and 2021 {#sec-Age}

It is crucial to analyze the distribution of arrests across different age groups to understand potential patterns and disparities. [@tbl-table1] presents a summary of arrest counts for various ethnic groups across all age categories in 2020 and 2021. The age groups included in the dataset are: Under 17, 18 to 24 years, 25 to 34 years, 35 to 44 years, 45 to 54 years, 55 to 64 years, and 65 years and over.

Given the complexity and extensive nature of the dataset, which involves nine age groups across eight ethnicities, summarizing the entire dataset in a comprehensible manner is challenging. To address this and ensure clarity, we have selected a subset of the data focusing specifically on individuals aged 25 to 34 years. This group was chosen as it represents a significant proportion of the data and allows for a more straightforward comparison of arrest trends across ethnic groups while avoiding information overload. By narrowing the scope, the analysis becomes more manageable and the insights more focused.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-table1
#| tbl-cap: "Arrests Booked by ethinic background for the year 2020, and 2021"

table3 <- cleaned_data %>%
  filter(!(age_group == "None") & !(perceived_race == "None")) |>
  group_by(arrest_year, perceived_race, age_group) %>%
  summarise(n = n(), .groups = "drop") %>%
  arrange(desc(perceived_race)) %>%
  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)

table3_small <- cleaned_data %>%
  filter(age_group == "Aged 25 to 34 years") |>
  group_by(arrest_year, perceived_race, age_group) %>%
  summarise(n = n(), .groups = "drop") |>
  arrange(desc(n)) |>
  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)

kable(table3_small,
      align = "c",
      linesep = "",
      col.names = c("Ethinic and Identity Based Groups", "Age Group", "2020", "2021"),
      booktabs = TRUE)

```

## Arrests by Ethnic and Identity-Based Groups for Year 2020 and 2021 {#sec-eth}

Our analysis primarily focuses on the distribution of arrests across different ethnic identities. We extracted data for the following ethnic groups: Black, White, South Asian, East Asian, Middle Eastern, Latin American, Southeast Asian, and Indigenous.

The data for 2020 and 2021 is summarized in [@tbl-table2] and [@tbl-table3]. [@tbl-table2] presents the number of individuals arrested from each ethnic group during these years, while [@tbl-table3] details how many of those arrested were subsequently booked under criminal charges.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-table2
#| tbl-cap: "Year 2020 vs Year 2021 Total number of arrests"

table1 <- cleaned_data |> 
  filter(!perceived_race == "None") |>
  group_by(arrest_year, perceived_race) |> 
  summarise(n = n(), .groups = "drop") |> 
  arrange(desc(n)) |> 
  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)

kable(table1,
      align = "c",
      linesep = "",
      col.names = c("Race", "2020", "2021"),
      booktabs = TRUE)
```

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-table3
#| tbl-cap: "Year 2020 vs Year 2021 Total number of Booked"

table2 <- cleaned_data |>
  filter(booked == 1) |>
  filter(!perceived_race == "None") |>
  group_by(arrest_year, perceived_race) |>
  summarise(n = n(), .groups = "drop") |>
  arrange(desc(n)) |>
  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)

kable(table2,
      align = "c",
      linesep = "",
      col.names = c("Race", "2020", "2021"),
      booktabs = TRUE)

```

It is important to note that the OpenDataToronto [@rOpenDataToronto] provides population distribution data by ethnicity. However, due to the size and diversity of Toronto and its numerous wards, it has been challenging to cross-verify this data with crime rate statistics in a comprehensive manner. Further, a more detailed analysis could be conducted by comparing specific ethnic groups directly.

Now that the data has been cleaned, we can begin to analyze the data. The following section will provide a summary of the data, and the trends that emerge from the data. [@fig-fig1], [@fig-fig2], and [@fig-fig3], was created using `ggplot` [@rGgplot2], displays the information that can be used as compare the trends as described above.

## Total Arrests in 2020-2021

Referring to [@tbl-table2], [@tbl-table3] and [@fig-fig1] the data reveals a non-uniform distribution of arrests across ethnic groups, with the Black and White populations showing the highest numbers of arrests in both 2020 and 2021. Additionally, the data indicates a decline in arrests for these two ethnic groups from 2020 to 2021. In contrast, arrests for other ethnic groups, such as South Asian, East Asian, and Indigenous populations, increased over the same period. These opposing trends highlight the divergent patterns of arrest across different ethnic groups during this time.

```{r}
#| warning: false
#| message: false
#| echo: false
#| eval: true
#| fig.cap: "# arrests by Ethnicity in Year 2020 and 2021"
#| fig.width: 10
#| fig.height: 6
#| fig.align: 'center'
#| fig.pos: 'htb'
#| label: fig-fig1

table1_long <- pivot_longer(table1, cols = c("2020", "2021"), names_to = "year", values_to = "value")

# Create a plot
ggplot(table1_long, aes(x = perceived_race, y = value, fill = year)) +
  geom_bar(stat = "identity", position = 'dodge') +
  labs(x = "Ethnicity",
       y = "# Arrests",
       fill = "Year") +
  theme_minimal() +
  scale_fill_discrete(labels = c("2020", "2021")) +
  theme(legend.position = "right") + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

Referring to [@tbl-table1] for the proportion of arrests that resulted in bookings, [@fig-fig2] was plotted with the y-axis representing the percentage of individuals booked after arrest. The data demonstrates a consistent distribution across various ethnic groups, with significantly more individuals being booked following arrest than those who were not. Notably, the percentage of individuals booked ranges between 50% and 55% across all ethnic groups, a pattern that remains consistent in 2021.

However, it is worth highlighting that the Black population had the highest number of individuals who were booked after arrest in both 2020 and 2021.

```{r}
#| warning: false
#| message: false
#| echo: false
#| eval: true
#| fig.cap: "Booked after Arrest in the Year 2020 and 2021"
#| fig.width: 10
#| fig.height: 6
#| fig.align: 'center'
#| fig.pos: 'htb'
#| label: fig-fig2

table2_long <-
  cleaned_data |>
  select(
    arrest_year,
    perceived_race,
    booked
  ) |>
  group_by(arrest_year, perceived_race, booked) %>%
  summarise(n = n(), .groups = "drop") %>%
  arrange(desc(perceived_race)) %>%
  filter(!(perceived_race == "None")) %>%
  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0)

table2_long <- pivot_longer(table2_long, cols = c("2020", "2021"), names_to = "year", values_to = "value")

ggplot(table2_long, aes(x = perceived_race, y = value, fill = factor(booked))) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~ year, scales = "free_y") +
  labs(x = "Perceived Race",
       y = "Percentage of Arrests Booked",
       fill = "Status") +
  theme_minimal() +
  theme(legend.position = "right") + 
  scale_fill_manual(values = c("0" = "lightblue", "1" = "darkblue"),
                    labels = c("0" = "Not Booked", "1" = "Booked")) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## Number of Arrests by Age Group for the year 2020-2021 {#sec-group}

An analysis of arrests by age groups is essential to understanding the demographic patterns of law enforcement activity. [@fig-fig3] illustrates the distribution of arrests across different age groups for the years 2020 and 2021. The age groups included in the dataset are Under 17, 18 to 24 years, 25 to 34 years, 35 to 44 years, 45 to 54 years, 55 to 64 years, and 65 years and over.

The data reveals a non-uniform distribution of arrests across age groups, with 25 to 34 and 35 to 44 age brackets accounting for the highest number of arrests in both 2020 and 2021. However, there is a notable decrease in the number of arrests for these age groups of 2020 to 2021 which indicates that there may be a possible shift in law enforcement activity.

In contrast, the other age groups experienced an increase in arrests during the same period. In particular, the number of arrests for individuals aged 18 to 24 and 45 to 54 saw an upward trend from 2020 to 2021.

```{r}
#| warning: false
#| message: false
#| echo: false
#| eval: true
#| fig.cap: "Number of arrests by Age-Group for the year 2020, and 2021"
#| fig.width: 10
#| fig.height: 6
#| fig.align: 'center'
#| fig.pos: 'htb'
#| label: fig-fig3
# Bar plot

table3_long <- pivot_longer(table3, cols = c("2020", "2021"), names_to = "year", values_to = "value")

ggplot(table3_long, aes(x = perceived_race, y = value, fill = factor(age_group))) +
  geom_bar(stat = "identity") +
  facet_wrap(~ year, scales = "free_y") +
  labs(x = "Perceived Race",
       y = "Number of People",
       fill = "Year") +
  theme_minimal() +
  theme(legend.position = "bottom") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Model {#sec-model}

This section describes the statistical model used to analyze the likelihood of an individual being booked after arrest, focusing on the relationship between ethnicity, year, and booking probabilities. A Bayesian logistic regression model was chosen due to its suitability for binary outcomes and its ability to quantify uncertainty in parameter estimates.

## Model Setup

\begin{align}
y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \label{eq1} \\
\mbox{logit}(\pi_i) &= \beta_0 + \beta_1 \times \mbox{race}_i + \beta_2 \times \mbox{year}_i \label{eq2} \\
\beta_0 &\sim \mbox{Normal}(0, 5) \label{eq3} \\
\beta_1 &\sim \mbox{Normal}(0, 5) \label{eq4} \\
\beta_2 &\sim \mbox{Normal}(0, 5) \label{eq5}
\end{align}

The Bayesian logistic regression model is represented by Equations (\ref{eq1}), (\ref{eq2}), (\ref{eq3}), (\ref{eq4}), and (\ref{eq5}). In these equations: - $y_i$ represents the booking status of the (i)-th individual (0 for not booked and 1 for booked). - $\pi_i$ represents the probability of the (i)-th individual being booked after arrest. - (\beta\_0) represents the intercept, or the log-odds of being booked for a White individual in the year 2020 (reference group). - (\beta\_1) represents how much the log-odds of being booked change on average depending on the race of the individual (relative to a White individual). - (\text{race}\_i) represents the race of the (i)-th individual (categorical predictor with levels: White, Black, Indigenous, South Asian, East/Southeast Asian, Middle Eastern, Latino, and Unknown/Legacy). - (\beta\_2) represents how much the log-odds of being booked change on average depending on the year (relative to 2020). - (\text{year}\_i) represents the year of the (i)-th individual’s arrest (binary: 0 for 2020, 1 for 2021).

The priors for (\beta\_0), (\beta\_1), and (\beta\_2) are weakly informative normal distributions with a mean of 0 and a standard deviation of 5. These priors allow for flexibility in parameter estimation while regularizing extreme values.

The (\beta) coefficients are important because they provide insight into how an individual’s race and the year of arrest influence the likelihood of being booked after arrest.

## Model Justification

The Bayesian logistic regression model was chosen for this analysis due to its suitability for modeling binary outcomes, its flexibility in handling uncertainty, and its ability to incorporate prior information. Below are the specific justifications for each aspect of the model design.

Logistic regression is the standard method for modeling binary response variables. In this case, the outcome of interest is whether an individual is booked after arrest (1 for booked, 0 for not booked). The logistic function ensures that the predicted probabilities fall between 0 and 1, making it ideal for this type of analysis. The log-odds output from logistic regression can be easily transformed into probabilities, providing an intuitive interpretation of the effects of predictors (e.g., how ethnicity or year influences the likelihood of booking). Logistic regression is well-suited for datasets where the outcome classes are imbalanced, as in this case, where the majority of individuals in most ethnic groups are booked.

The Bayesian approach explicitly models uncertainty in the parameter estimates through posterior distributions. This is particularly important in cases like this, where: Data for some racial groups (e.g., Indigenous, East/Southeast Asian) are sparse. Estimating small differences in booking probabilities requires careful uncertainty assessment. Bayesian models provide credibility intervals, which are more intuitive than traditional frequentist confidence intervals. A 90% credibility interval directly indicates the range within which the parameter value lies with 90% probability, given the data and prior assumptions.

## Why Bayesian Logistic Regression Over Alternatives?

Linear regression was not used because it is unsuitable for binary outcomes, as it can produce predictions outside the range of 0 to 1. Additionally, the relationship between the predictors and the outcome is not linear in this context, violating the assumptions of linear regression.

Frequentist logistic regression does not account for prior knowledge or explicitly quantify uncertainty in parameter estimates. Given the small sample sizes for some groups, a Bayesian approach is more robust and transparent.

The model was designed to provide insights into: Whether racial groups differ in booking probabilities after arrest. Whether booking probabilities changed between 2020 and 2021. The Bayesian framework ensures these insights are backed by a robust statistical foundation that accounts for both data-driven and prior uncertainties.

This justification underscores the appropriateness of the Bayesian logistic regression model for analyzing disparities in booking probabilities while addressing potential limitations of the data and context. It aligns the methodology with the study’s objectives, ensuring clarity and rigor in the analysis.

# Results {#sec-results}

This section presents the findings of the Bayesian logistic regression model used to analyze booking probabilities across ethnic groups and between the years 2020 and 2021. The results focus on the coefficients of the model, posterior predictive checks, and the observed trends in booking probabilities.

## Model Outputs

```{r}
#| echo: false
#| eval: true
#| warning: false
#| label: tbl-model
#| tbl-cap: "Modeling the likelihood arrested person getting booked"

modelsummary(
  list(
    "Booked on Racial Profile" = model
  ),
  statistic = "mad"
)
```

[@tbl-model] estimates the log-odds of being booked after arrest as a function of ethnicity and year. The coefficients, their posterior means, and the associated uncertainty (standard deviations and credibility intervals) are summarized below:

Intercept −8.859 The intercept represents the log-odds of being booked for the reference group (White individuals in 2020). A large negative value indicates a low baseline probability of booking for this group.

Ethnicity Effects Coefficients for all racial groups (e.g., Black, Indigenous, South Asian) relative to the reference group (White) were close to 0. Credibility intervals were wide, reflecting high uncertainty in these estimates. For example: perceived_raceBlack: Mean = -0.025, SD = 4.236 perceived_raceIndigenous: Mean = 0.118, SD = 4.240 These results suggest no strong evidence of systematic differences in booking probabilities across ethnic groups after arrest, but high variability in estimates limits definitive conclusions.

Year Effect −0.076 This indicates a slight decline in booking probabilities compared to 2020, but the effect is negligible and statistically uncertain (high variability with SD = 3.466). This suggests booking practices remained relatively stable between the two years.

## Posterior Predictive Checks

Posterior predictive checks were conducted to evaluate the model’s ability to replicate observed data. Key observations:

The model accurately predicted the central trend in booking probabilities (\~50% for most groups). It struggled to capture variability across ethnic groups and years due to high uncertainty in some coefficients. The predictive checks indicated a reasonable overall fit, but with potential underfitting for smaller groups like Indigenous and East/Southeast Asian populations.

## Observed Patterns in Booking Probabilities

Ethnic Groups:

Booking probabilities ranged consistently between 50% and 55% for most ethnic groups in both years. While Black individuals had the highest number of arrests and bookings, their booking probability was not significantly different from other groups after controlling for ethnicity and year. Year-to-Year Trends:

Booking probabilities were stable across years, with minor declines observed for certain groups (e.g., White, Black) and slight increases for others (e.g., Indigenous, Middle Eastern). These trends were not statistically significant within the model. Uncertainty in Estimates:

Wide credibility intervals for racial group coefficients suggest that data limitations (e.g., small sample sizes for some groups) contributed to high variability in results. This highlights the need for more comprehensive data to draw robust conclusions.

# Discussion {#sec-discussion}

This report analyzed arrest and booking data for Toronto during 2020 and 2021 using a Bayesian logistic regression model to examine predictors of booking probabilities. The predictors in this analysis were ethnicity and year. The results provide insights into equity in booking practices and the implications of these findings for law enforcement policies.

## Black and Indigenous individuals are overrepresented in arrests but not in booking probabilities

Analysis of arrest data revealed that Black and Indigenous individuals were overrepresented in arrests relative to their share of the population, consistent with prior findings from the Toronto Police Service and other studies on racial disparities in policing. However, this study focused on booking probabilities after arrest and found no significant differences in booking outcomes across ethnic groups after controlling for year.

The wide credibility intervals in the Bayesian model, especially for smaller groups like Indigenous and East/Southeast Asian individuals, reflect high uncertainty. While this limits definitive conclusions about booking disparities, it suggests that any disparities observed in arrests may not persist at the booking stage. However, data limitations and measurement biases—such as officer-recorded perceived race—may influence these findings.

## Booking probabilities remained stable across 2020 and 2021

The model found no significant change in booking probabilities between 2020 and 2021. The year coefficient (-0.076) was small and statistically insignificant, indicating stability in booking practices over the two years. This result suggests that broader procedural norms or operational policies likely remained consistent during this period, despite heightened societal focus on police reform.

The stability in booking probabilities aligns with the broader literature on procedural safeguards in booking decisions. While arrests may be influenced by discretionary or situational factors, booking practices often involve formalized procedures that may reduce variability. However, the absence of contextual variables in this dataset, such as offense type or search status, limits the ability to fully explain the observed stability.

## Implications for policy and public discourse

The absence of strong evidence for racial disparities in booking probabilities is encouraging for equitable law enforcement practices at this stage. However, the high uncertainty in estimates for smaller groups highlights the need for continued monitoring and data improvements. Policymakers should focus on enhancing data collection practices and transparency in booking decisions while considering potential disparities earlier (e.g., arrests) or later (e.g., sentencing) in the justice system.

Communicating these findings is critical for informed public discourse. While arrest data often highlights disparities, the results suggest that booking probabilities may not exhibit the same biases. This distinction can help guide discussions on police reform and support targeted interventions to address inequities at all stages of law enforcement decision-making.

## Limitations and future research

### Data Limitations

The dataset relied on officer-recorded perceived race and binary booking outcomes, omitting variables such as offense type, search status, or socio-economic indicators. These omissions may obscure nuanced influences on booking decisions and contribute to the high uncertainty in coefficient estimates for smaller groups.

## Future Research

1.  **Incorporate Contextual Variables**: Adding variables like offense type or search status would provide a more comprehensive view of factors influencing booking probabilities.
2.  **Examine Interaction Effects**: Analyzing interactions between race and year could highlight evolving trends for specific groups.
3.  **Use Hierarchical Models**: Regional analyses could identify precinct-level disparities or best practices in booking decisions.
4.  **Explore Arrest-to-Booking Pipelines**: Investigating disparities across the entire law enforcement process—from arrests to sentencing—would provide a holistic understanding of equity in policing.

# Appendix

## Model details {#sec-model-details}

### Posterior predictive check and comparison of the prior and posterior {#sec-ppc-comparison}

(@fig-ppc) shows posterior predictive check was performed . @fig-ppc illustrates a comparison between posterior distribution simulations and the actual probability of experiencing a custody arrest

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppc
#| fig-cap: "Posterior predictive check"

pp_check(model) +
  theme_classic() +
  theme(legend.position = "bottom")
```

In addition, using code adapted from @tellingstorieswithdata, a comparison of the prior and posterior was performed (@fig-comparison). @fig-comparison depicts the amounts by which the estimates are modified after considering data [@tellingstorieswithdata].

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-comparison
#| fig-cap: "Comparison of the prior and posterior"

# The code below was adapted based on Alexander (2023).
# 
# posterior_vs_prior(model) +
#   theme_minimal() +
#   scale_color_brewer(palette = "Set1") +
#   theme(legend.position = "none") +
#   coord_flip()
```

### Diagnostics {#sec-diagnostics}

This section assesses whether the `rstanarm` [@citerstanarm] Markov chain Monte Carlo (MCMC) algorithm experienced problems [@tellingstorieswithdata].

Using code adapted from @tellingstorieswithdata, a trace plot was generated (@fig-trace). @fig-trace illustrates horizontal lines that are seemingly erratic with chain overlap, indicating nothing unusual (based on @tellingstorieswithdata).

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-trace
#| fig-cap: "Trace plot"

plot(model, "trace")
```

Moreover, using code adapted from @tellingstorieswithdata, a Rhat plot was created (@fig-rhat). @fig-rhat depicts values near 1 and less than or equal to 1.1, indicating no issues (based on @tellingstorieswithdata).

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-rhat
#| fig-cap: "Rhat plot"

# The code below was adapted from Alexander (2023).

plot(model, "rhat")
```

## Credibility intervals {#sec-credibility-intervals}

Using code adapted from @tellingstorieswithdata, @fig-credibility-intervals-all was generated. @fig-credibility-intervals-all illustrates 90\% credibility intervals of all of the predictors of undergoing a custody arrest based on the model from this report.

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-credibility-intervals-all
#| fig-cap: "Credibility intervals of all of the custody arrest predictors"

# The code below was adapted from Alexander (2023).

modelplot(model, conf_level = 0.9) +
  labs(x = "90 percent credibility interval")
```

@fig-credibility-intervals-several depicts 90\% credibility intervals of several of the predictors of undergoing a custody arrest. In particular, the American Indian/Alaskan Native and Hawaiian/Pacific Islander predictors were omitted for visual clarity (as compared to @fig-credibility-intervals-all).

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-credibility-intervals-several
#| fig-cap: "Credibility intervals of several of the custody arrest predictors"

# Portions of the code below were adapted from Alexander (2023).

modelplot(model, conf_level = 0.9, coef_omit = c(5, 8)) +
  labs(x = "90 percent credibility interval")

```

## Observational Data and Sampling in Law Enforcement Studies

### Observational Data Context

The dataset reflects arrest and booking events recorded by law enforcement in Toronto during 2020 and 2021. As observational data, it represents real-world interactions but is subject to biases like selection bias (e.g., only arrested individuals are included) and measurement bias (e.g., officer-perceived race).

### Sampling Bias

The dataset captures only individuals who were arrested, excluding those who interacted with law enforcement but were not arrested. This selection bias may affect the generalizability of the findings to all police interactions.

### Measurement Bias

Key variables, such as `perceived_race`, are officer-reported and subject to subjective errors. Misclassification in smaller groups (e.g., Indigenous or East/Southeast Asian) can inflate or deflate booking probabilities for these groups, distorting the analysis. This aligns with literature emphasizing the need for objective self-reported data in law enforcement studies.

### Implications and Recommendations

1.  **Expand Variables**: Incorporating variables like offense type or socio-economic indicators can address confounding.
2.  **Account for Bias**: Sensitivity analyses can estimate the impact of misclassification or selection bias.
3.  **Improve Data Practices**: Encouraging self-reported race data and capturing broader police interactions can enhance representativeness.

These considerations highlight the strengths and limitations of observational datasets for analyzing law enforcement practices, emphasizing the importance of robust methods to ensure accurate interpretations.

\newpage

# References

### 2024-12-03 Bias-in-arrests-toronto/scripts/01-test_simulated_data.R

@@ -1,0 +1,123 @@
+#### Preamble ####
+# Purpose: Downloading and saving the data from OpenData Toronto for Police Race and Identity Based Data Collection: Arrests & Strip Searches
+# Author: Akshat Aneja
+# Date: 1 December 2024
+# Contact: akshat.aneja@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: none
+# Datasets: https://open.toronto.ca/dataset/police-race-and-identity-based-data-collection-arrests-strip-searches/
+
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(testthat)
+
+analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv", show_col_types = FALSE)
+
+
+# Define the tests
+
+# Test 1: Dataset is a data frame
+test_that("Dataset is a data frame", {
+  expect_true(is.data.frame(simulated_data))
+})
+
+# Test 2: Number of columns is correct
+test_that("Number of columns is correct", {
+  expect_equal(ncol(simulated_data), 6)
+})
+
+# Test 3: Column names are correct
+test_that("Column names are correct", {
+  expect_equal(colnames(simulated_data), 
+               c("arrest_year", "age_group", "booked", 
+                 "search_reason", "perceived_race", "minor"))
+})
+
+# Test 4: `arrest_year` contains only 2020 or 2021
+test_that("`arrest_year` contains only 2020 or 2021", {
+  expect_true(all(simulated_data$arrest_year %in% c(2020, 2021)))
+})
+
+# Test 5: `age_group` contains only valid categories
+test_that("`age_group` contains only valid categories", {
+  valid_age_groups <- c("Aged 18 to 24 years", "Aged 25 to 34 years", 
+                        "Aged 35 to 44 years", "Aged 45 to 54 years", 
+                        "Aged 55 years and older")
+  expect_true(all(simulated_data$age_group %in% valid_age_groups))
+})
+
+# Test 6: `booked` contains only 0 or 1
+test_that("`booked` contains only 0 or 1", {
+  expect_true(all(simulated_data$booked %in% c(0, 1)))
+})
+
+# Test 7: `search_reason` contains only 0 or 1
+test_that("`search_reason` contains only 0 or 1", {
+  expect_true(all(simulated_data$search_reason %in% c(0, 1)))
+})
+
+# Test 8: `perceived_race` contains only valid categories
+test_that("`perceived_race` contains only valid categories", {
[…]

### 2024-12-03 Bias-in-arrests-toronto/scripts/02-download_data.R

@@ -1,0 +1,34 @@
+#### Preamble ####
+# Purpose: Downloading and saving the data from OpenData Toronto for Police Race and Identity Based Data Collection: Arrests & Strip Searches
+# Author: Akshat Aneja
+# Date: 23 September 2024
+# Contact: akshat.aneja@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: none
+# Datasets: https://open.toronto.ca/dataset/police-race-and-identity-based-data-collection-arrests-strip-searches/
+
+
+
+#### Workspace setup ####
+library(opendatatoronto)
+library(tidyverse)
+# Get package data
+
+# get package
+package <- show_package("police-race-and-identity-based-data-collection-arrests-strip-searches")
+package
+
+# get all resources for this package
+resources <- list_package_resources("police-race-and-identity-based-data-collection-arrests-strip-searches")
+
+# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+
+# load the first datastore resource as a sample
+data <- filter(datastore_resources, row_number()==1) %>% get_resource()
+
+# Save the data in CSV
+write.csv(data, here::here("data/01-raw_data/raw_data.csv"), row.names = FALSE)
+
+
+         

### 2024-12-03 Bias-in-arrests-toronto/scripts/03-clean_data.R

@@ -1,0 +1,69 @@
+#### Preamble ####
+# Purpose: Downloading and saving the data from OpenData Toronto for Police Race and Identity Based Data Collection: Arrests & Strip Searches
+# Author: Akshat Aneja
+# Date: 23 September 2024
+# Contact: akshat.aneja@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: none
+# Datasets: https://open.toronto.ca/dataset/police-race-and-identity-based-data-collection-arrests-strip-searches/
+
+#### Workspace setup ####
+library(tidyverse)
+library(arrow)
+
+#### Clean data ####
+data <- read.csv(here::here("data/01-raw_data/raw_data.csv"))
+
+cleaned_data <- data |>
+  select(
+    Arrest_Year, Perceived_Race, Age_group__at_arrest_, Youth_at_arrest__under_18_years, Booked, SearchReason_CauseInjury, SearchReason_AssistEscape, SearchReason_PossessWeapons, SearchReason_PossessEvidence)
+
+cleaned_data <- cleaned_data %>%
+  mutate(search_reason = case_when(
+    SearchReason_CauseInjury == 1 ~ "1",
+    SearchReason_AssistEscape == 1 ~ "1",
+    SearchReason_PossessWeapons == 1 ~ "1",
+    SearchReason_PossessEvidence == 1 ~ "1",
+    TRUE ~ "0"
+  ))
+
+cleaned_data <- cleaned_data %>%
+  rename(
+    age_group = Age_group__at_arrest_,
+    youth_at_arrest_under_18_years = Youth_at_arrest__under_18_years
+  )
+
+cleaned_data <- cleaned_data %>%
+  mutate(age_group = case_when(
+    age_group == "Aged 17 years and younger" ~ "Under 17",
+    age_group == "Aged 17 years and under" ~ "Under 17",
+    TRUE ~ age_group
+  ))
+
+cleaned_data <- cleaned_data %>%
+  mutate(booked = case_when(
+    is.na(Booked) ~ "0",
+    Booked == 1 ~ "1",
+    TRUE ~ "Not Booked"
+  ))
+
+#### Cleaned Data Conversion to Numeric ####
+
+# Convert arrest_year, search_reason, and booked to numeric
+cleaned_data <- cleaned_data |>
+  select(
+    Arrest_Year, age_group, booked, search_reason, Perceived_Race, youth_at_arrest_under_18_years) |>
+  rename(
[…]

### 2024-12-03 Bias-in-arrests-toronto/scripts/06-model_data.R

@@ -1,0 +1,73 @@
+#### Preamble ####
+# Purpose: Models... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+
+#### Read data ####
+
+cleaned_data = read.csv(here::here("data/02-analysis_data/analysis_data.csv"))
+
+arrest_data <-
+  cleaned_data |>
+  filter(!perceived_race == "None") |>
+  group_by(arrest_year, perceived_race) |> 
+  summarise(n = n(), .groups = "drop") |> 
+  arrange(desc(n)) |> 
+  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0) |>
+  rename(arrests_2020 = '2020', arrests_2021 = '2021')
+
+booked_data <-
+  cleaned_data |>
+  filter(booked == 1) |>
+  filter(!perceived_race == "None") |>
+  group_by(arrest_year, perceived_race) |>
+  summarise(n = n(), .groups = "drop") |>
+  arrange(desc(n)) |>
+  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0) |>
+  rename(booked_2020 = '2020', booked_2021 = '2021')
+
+combined_data <- merge(arrest_data, booked_data, by = "perceived_race")
+
+# Calculate booking probability for each year
+combined_data$prob_2020 <- combined_data$booked_2020 / combined_data$arrests_2020
+combined_data$prob_2021 <- combined_data$booked_2021 / combined_data$arrests_2021
+
+# Reshape the data into a long format
+long_data <- combined_data %>%
+  pivot_longer(cols = c(prob_2020, prob_2021),
+               names_to = "year",
+               values_to = "booking_prob")
+
+# Convert 'year' and 'perceived_race' to factors for the model
+long_data$year <- as.factor(long_data$year)  # Ensure 'year' is a factor
+long_data$perceived_race <- as.factor(long_data$perceived_race)  # Ensure 'perceived_race' is a factor
+
+### Model data ####
+# Fit the Bayesian logistic regression model for both years
+model <- stan_glm(
[…]

### 2024-12-03 scripts/00-simulate_data.R
Fixing file structure

@@ -1,0 +1,40 @@
+#### Preamble ####
+# Purpose: Downloading and saving the data from OpenData Toronto for Police Race and Identity Based Data Collection: Arrests & Strip Searches
+# Author: Akshat Aneja
+# Date: 23 September 2024
+# Contact: akshat.aneja@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: none
+# Datasets: https://open.toronto.ca/dataset/police-race-and-identity-based-data-collection-arrests-strip-searches/
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(dplyr)
+
+#### Simulate data ####
+set.seed(1007901157)
+
+# Define the number of rows to simulate
+n <- 300
+
+# Simulate the data
+simulated_data <- data.frame(
+  arrest_year = sample(c(2020, 2021), n, replace = TRUE),
+  age_group = sample(c("Aged 18 to 24 years", "Aged 25 to 34 years", 
+                       "Aged 35 to 44 years", "Aged 45 to 54 years", 
+                       "Aged 55 years and older"), n, replace = TRUE),
+  booked = sample(c(0, 1), n, replace = TRUE),
+  search_reason = sample(0:1, n, replace = TRUE),
+  perceived_race = sample(c("White", "Black", "South Asian", 
+                            "East Asian", "Hispanic", "Unknown or Legacy"), n, replace = TRUE),
+  minor = sample(c("Not a youth", "Youth"), n, replace = TRUE)
+)
+
+
+## Write the simulated data
+write.csv(simulated_data, here::here("data/00-simulated_data/simulated_data.csv"), row.names = FALSE)
+
+
+
+

### 2024-12-03 scripts/04-test_analysis_data.R

@@ -1,0 +1,96 @@
+#### Preamble ####
+# Purpose: Downloading and saving the data from OpenData Toronto for Police Race and Identity Based Data Collection: Arrests & Strip Searches
+# Author: Akshat Aneja
+# Date: 23 September 2024
+# Contact: akshat.aneja@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: none
+# Datasets: https://open.toronto.ca/dataset/police-race-and-identity-based-data-collection-arrests-strip-searches/
+
+
+#### Workspace setup ####
+library(tidyverse)
+
+
+#### Test data ####
+arrest_data <- read.csv(here::here("data/02-analysis_data/analysis_data.csv"))
+
+# 1. Check if 'arrest_year' is numeric
+is_arrest_year_numeric <- all(as.numeric(arrest_data$arrest_year) == arrest_data$arrest_year)
+
+# 2. Validate that 'booked' is numeric
+is_booked_numeric <- all(as.numeric(arrest_data$booked) == arrest_data$booked)
+
+# 3. Validate that 'search_reason' is numeric
+is_search_reason_numeric <- all(as.numeric(arrest_data$search_reason) == arrest_data$search_reason)
+
+# 4. Check if 'age_group' is of class character
+is_age_group_character <- class(arrest_data$age_group) == "character"
+
+# 5. Check if 'perceived_race' is of class character
+is_perceived_race_character <- class(arrest_data$perceived_race) == "character"
+
+# 6. Check if 'minor' is of class character
+is_minor_character <- class(arrest_data$minor) == "character"
+
+# 7. Validate that 'booked' values are between 0 and 1
+valid_booked_values <- all(arrest_data$booked >= 0 & arrest_data$booked <= 1)
+
+# 8. Validate that 'arrest_year' values are in a reasonable range
+valid_arrest_year_values <- all(arrest_data$arrest_year >= 2020 & arrest_data$arrest_year <= 2021)
+
+# 9. Validate that 'search_reason' values are binary (0 or 1)
+valid_search_reason_values <- all(arrest_data$search_reason >= 0 & arrest_data$search_reason <= 1)
+
+
+
+# 12. Check for missing values in the dataset
+no_missing_values <- all(complete.cases(arrest_data))
+
+
+
+
+# 15. Validate 'booked' column as logical or numeric
+booked_logical <- is.logical(arrest_data$booked) || is.numeric(arrest_data$booked)
+
[…]

### 2024-12-03 Bias-in-arrests-toronto/scripts/00-simulate_data.R
Delete Bias-in-arrests-toronto directory

@@ -1,40 +1,0 @@
-#### Preamble ####
-# Purpose: Downloading and saving the data from OpenData Toronto for Police Race and Identity Based Data Collection: Arrests & Strip Searches
-# Author: Akshat Aneja
-# Date: 23 September 2024
-# Contact: akshat.aneja@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: none
-# Datasets: https://open.toronto.ca/dataset/police-race-and-identity-based-data-collection-arrests-strip-searches/
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(dplyr)
-
-#### Simulate data ####
-set.seed(1007901157)
-
-# Define the number of rows to simulate
-n <- 300
-
-# Simulate the data
-simulated_data <- data.frame(
-  arrest_year = sample(c(2020, 2021), n, replace = TRUE),
-  age_group = sample(c("Aged 18 to 24 years", "Aged 25 to 34 years", 
-                       "Aged 35 to 44 years", "Aged 45 to 54 years", 
-                       "Aged 55 years and older"), n, replace = TRUE),
-  booked = sample(c(0, 1), n, replace = TRUE),
-  search_reason = sample(0:1, n, replace = TRUE),
-  perceived_race = sample(c("White", "Black", "South Asian", 
-                            "East Asian", "Hispanic", "Unknown or Legacy"), n, replace = TRUE),
-  minor = sample(c("Not a youth", "Youth"), n, replace = TRUE)
-)
-
-
-## Write the simulated data
-write.csv(simulated_data, here::here("data/00-simulated_data/simulated_data.csv"), row.names = FALSE)
-
-
-
-

### 2024-12-03 Bias-in-arrests-toronto/scripts/01-test_simulated_data.R

@@ -1,123 +1,0 @@
-#### Preamble ####
-# Purpose: Downloading and saving the data from OpenData Toronto for Police Race and Identity Based Data Collection: Arrests & Strip Searches
-# Author: Akshat Aneja
-# Date: 1 December 2024
-# Contact: akshat.aneja@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: none
-# Datasets: https://open.toronto.ca/dataset/police-race-and-identity-based-data-collection-arrests-strip-searches/
-
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(testthat)
-
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv", show_col_types = FALSE)
-
-
-# Define the tests
-
-# Test 1: Dataset is a data frame
-test_that("Dataset is a data frame", {
-  expect_true(is.data.frame(simulated_data))
-})
-
-# Test 2: Number of columns is correct
-test_that("Number of columns is correct", {
-  expect_equal(ncol(simulated_data), 6)
-})
-
-# Test 3: Column names are correct
-test_that("Column names are correct", {
-  expect_equal(colnames(simulated_data), 
-               c("arrest_year", "age_group", "booked", 
-                 "search_reason", "perceived_race", "minor"))
-})
-
-# Test 4: `arrest_year` contains only 2020 or 2021
-test_that("`arrest_year` contains only 2020 or 2021", {
-  expect_true(all(simulated_data$arrest_year %in% c(2020, 2021)))
-})
-
-# Test 5: `age_group` contains only valid categories
-test_that("`age_group` contains only valid categories", {
-  valid_age_groups <- c("Aged 18 to 24 years", "Aged 25 to 34 years", 
-                        "Aged 35 to 44 years", "Aged 45 to 54 years", 
-                        "Aged 55 years and older")
-  expect_true(all(simulated_data$age_group %in% valid_age_groups))
-})
-
-# Test 6: `booked` contains only 0 or 1
-test_that("`booked` contains only 0 or 1", {
-  expect_true(all(simulated_data$booked %in% c(0, 1)))
-})
-
-# Test 7: `search_reason` contains only 0 or 1
-test_that("`search_reason` contains only 0 or 1", {
-  expect_true(all(simulated_data$search_reason %in% c(0, 1)))
-})
-
-# Test 8: `perceived_race` contains only valid categories
-test_that("`perceived_race` contains only valid categories", {
[…]

### 2024-12-03 Bias-in-arrests-toronto/scripts/02-download_data.R

@@ -1,34 +1,0 @@
-#### Preamble ####
-# Purpose: Downloading and saving the data from OpenData Toronto for Police Race and Identity Based Data Collection: Arrests & Strip Searches
-# Author: Akshat Aneja
-# Date: 23 September 2024
-# Contact: akshat.aneja@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: none
-# Datasets: https://open.toronto.ca/dataset/police-race-and-identity-based-data-collection-arrests-strip-searches/
-
-
-
-#### Workspace setup ####
-library(opendatatoronto)
-library(tidyverse)
-# Get package data
-
-# get package
-package <- show_package("police-race-and-identity-based-data-collection-arrests-strip-searches")
-package
-
-# get all resources for this package
-resources <- list_package_resources("police-race-and-identity-based-data-collection-arrests-strip-searches")
-
-# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
-datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
-
-# load the first datastore resource as a sample
-data <- filter(datastore_resources, row_number()==1) %>% get_resource()
-
-# Save the data in CSV
-write.csv(data, here::here("data/01-raw_data/raw_data.csv"), row.names = FALSE)
-
-
-         

### 2024-12-03 Bias-in-arrests-toronto/scripts/03-clean_data.R

@@ -1,69 +1,0 @@
-#### Preamble ####
-# Purpose: Downloading and saving the data from OpenData Toronto for Police Race and Identity Based Data Collection: Arrests & Strip Searches
-# Author: Akshat Aneja
-# Date: 23 September 2024
-# Contact: akshat.aneja@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: none
-# Datasets: https://open.toronto.ca/dataset/police-race-and-identity-based-data-collection-arrests-strip-searches/
-
-#### Workspace setup ####
-library(tidyverse)
-library(arrow)
-
-#### Clean data ####
-data <- read.csv(here::here("data/01-raw_data/raw_data.csv"))
-
-cleaned_data <- data |>
-  select(
-    Arrest_Year, Perceived_Race, Age_group__at_arrest_, Youth_at_arrest__under_18_years, Booked, SearchReason_CauseInjury, SearchReason_AssistEscape, SearchReason_PossessWeapons, SearchReason_PossessEvidence)
-
-cleaned_data <- cleaned_data %>%
-  mutate(search_reason = case_when(
-    SearchReason_CauseInjury == 1 ~ "1",
-    SearchReason_AssistEscape == 1 ~ "1",
-    SearchReason_PossessWeapons == 1 ~ "1",
-    SearchReason_PossessEvidence == 1 ~ "1",
-    TRUE ~ "0"
-  ))
-
-cleaned_data <- cleaned_data %>%
-  rename(
-    age_group = Age_group__at_arrest_,
-    youth_at_arrest_under_18_years = Youth_at_arrest__under_18_years
-  )
-
-cleaned_data <- cleaned_data %>%
-  mutate(age_group = case_when(
-    age_group == "Aged 17 years and younger" ~ "Under 17",
-    age_group == "Aged 17 years and under" ~ "Under 17",
-    TRUE ~ age_group
-  ))
-
-cleaned_data <- cleaned_data %>%
-  mutate(booked = case_when(
-    is.na(Booked) ~ "0",
-    Booked == 1 ~ "1",
-    TRUE ~ "Not Booked"
-  ))
-
-#### Cleaned Data Conversion to Numeric ####
-
-# Convert arrest_year, search_reason, and booked to numeric
-cleaned_data <- cleaned_data |>
-  select(
-    Arrest_Year, age_group, booked, search_reason, Perceived_Race, youth_at_arrest_under_18_years) |>
-  rename(
[…]

### 2024-12-03 Bias-in-arrests-toronto/scripts/04-test_analysis_data.R

@@ -1,96 +1,0 @@
-#### Preamble ####
-# Purpose: Downloading and saving the data from OpenData Toronto for Police Race and Identity Based Data Collection: Arrests & Strip Searches
-# Author: Akshat Aneja
-# Date: 23 September 2024
-# Contact: akshat.aneja@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: none
-# Datasets: https://open.toronto.ca/dataset/police-race-and-identity-based-data-collection-arrests-strip-searches/
-
-
-#### Workspace setup ####
-library(tidyverse)
-
-
-#### Test data ####
-arrest_data <- read.csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-# 1. Check if 'arrest_year' is numeric
-is_arrest_year_numeric <- all(as.numeric(arrest_data$arrest_year) == arrest_data$arrest_year)
-
-# 2. Validate that 'booked' is numeric
-is_booked_numeric <- all(as.numeric(arrest_data$booked) == arrest_data$booked)
-
-# 3. Validate that 'search_reason' is numeric
-is_search_reason_numeric <- all(as.numeric(arrest_data$search_reason) == arrest_data$search_reason)
-
-# 4. Check if 'age_group' is of class character
-is_age_group_character <- class(arrest_data$age_group) == "character"
-
-# 5. Check if 'perceived_race' is of class character
-is_perceived_race_character <- class(arrest_data$perceived_race) == "character"
-
-# 6. Check if 'minor' is of class character
-is_minor_character <- class(arrest_data$minor) == "character"
-
-# 7. Validate that 'booked' values are between 0 and 1
-valid_booked_values <- all(arrest_data$booked >= 0 & arrest_data$booked <= 1)
-
-# 8. Validate that 'arrest_year' values are in a reasonable range
-valid_arrest_year_values <- all(arrest_data$arrest_year >= 2020 & arrest_data$arrest_year <= 2021)
-
-# 9. Validate that 'search_reason' values are binary (0 or 1)
-valid_search_reason_values <- all(arrest_data$search_reason >= 0 & arrest_data$search_reason <= 1)
-
-
-
-# 12. Check for missing values in the dataset
-no_missing_values <- all(complete.cases(arrest_data))
-
-
-
-
-# 15. Validate 'booked' column as logical or numeric
-booked_logical <- is.logical(arrest_data$booked) || is.numeric(arrest_data$booked)
-
[…]

### 2024-12-03 Bias-in-arrests-toronto/scripts/06-model_data.R

@@ -1,73 +1,0 @@
-#### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(rstanarm)
-
-#### Read data ####
-
-cleaned_data = read.csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-arrest_data <-
-  cleaned_data |>
-  filter(!perceived_race == "None") |>
-  group_by(arrest_year, perceived_race) |> 
-  summarise(n = n(), .groups = "drop") |> 
-  arrange(desc(n)) |> 
-  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0) |>
-  rename(arrests_2020 = '2020', arrests_2021 = '2021')
-
-booked_data <-
-  cleaned_data |>
-  filter(booked == 1) |>
-  filter(!perceived_race == "None") |>
-  group_by(arrest_year, perceived_race) |>
-  summarise(n = n(), .groups = "drop") |>
-  arrange(desc(n)) |>
-  pivot_wider(names_from = arrest_year, values_from = n, values_fill = 0) |>
-  rename(booked_2020 = '2020', booked_2021 = '2021')
-
-combined_data <- merge(arrest_data, booked_data, by = "perceived_race")
-
-# Calculate booking probability for each year
-combined_data$prob_2020 <- combined_data$booked_2020 / combined_data$arrests_2020
-combined_data$prob_2021 <- combined_data$booked_2021 / combined_data$arrests_2021
-
-# Reshape the data into a long format
-long_data <- combined_data %>%
-  pivot_longer(cols = c(prob_2020, prob_2021),
-               names_to = "year",
-               values_to = "booking_prob")
-
-# Convert 'year' and 'perceived_race' to factors for the model
-long_data$year <- as.factor(long_data$year)  # Ensure 'year' is a factor
-long_data$perceived_race <- as.factor(long_data$perceived_race)  # Ensure 'perceived_race' is a factor
-
-### Model data ####
-# Fit the Bayesian logistic regression model for both years
-model <- stan_glm(
[…]

### 2024-12-03 scripts/00-simulate_data.R
Add files via upload

@@ -1,7 +1,7 @@
 #### Preamble ####
-# Purpose: Downloading and saving the data from OpenData Toronto for Police Race and Identity Based Data Collection: Arrests & Strip Searches
+# Purpose: Downloading and storing data from OpenData Toronto on Police Race and Identity-Based Data Collection: Arrests and Strip Searches.
 # Author: Akshat Aneja
-# Date: 23 September 2024
+# Date: 3 December 2024
 # Contact: akshat.aneja@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: none

### 2024-12-03 scripts/01-test_simulated_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
-# Purpose: Downloading and saving the data from OpenData Toronto for Police Race and Identity Based Data Collection: Arrests & Strip Searches
+# Purpose: Testing the simulated data from OpenData Toronto on Police Race and Identity-Based Data Collection: Arrests and Strip Searches.
 # Author: Akshat Aneja
-# Date: 1 December 2024
+# Date: 3 December 2024
 # Contact: akshat.aneja@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: none
@@ -15,8 +15,6 @@ library(testthat)
 
 analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv", show_col_types = FALSE)
 
-
-# Define the tests
 
 # Test 1: Dataset is a data frame
 test_that("Dataset is a data frame", {
@@ -75,48 +73,47 @@ test_that("No missing values in the dataset", {
   expect_true(all(!is.na(simulated_data)))
 })
 
-
-# Test 12: `arrest_year` is numeric
+# Test 11: `arrest_year` is numeric
 test_that("`arrest_year` is numeric", {
   expect_true(is.numeric(simulated_data$arrest_year))
 })
 
-# Test 13: `age_group` is a factor or character
+# Test 12: `age_group` is a factor or character
 test_that("`age_group` is a factor or character", {
   expect_true(is.character(simulated_data$age_group) || is.factor(simulated_data$age_group))
 })
 
-# Test 14: `booked` is numeric
+# Test 13: `booked` is numeric
 test_that("`booked` is numeric", {
   expect_true(is.numeric(simulated_data$booked))
 })
 
-# Test 15: `search_reason` is numeric
+# Test 14: `search_reason` is numeric
 test_that("`search_reason` is numeric", {
   expect_true(is.numeric(simulated_data$search_reason))
 })
 
-# Test 16: `perceived_race` is a factor or character
+# Test 15: `perceived_race` is a factor or character
 test_that("`perceived_race` is a factor or character", {
   expect_true(is.character(simulated_data$perceived_race) || is.factor(simulated_data$perceived_race))
 })
 
-# Test 17: `minor` is a factor or character
+# Test 16: `minor` is a factor or character
 test_that("`minor` is a factor or character", {
   expect_true(is.character(simulated_data$minor) || is.factor(simulated_data$minor))
 })
 
-# Test 18: `booked` contains at least one 0 and one 1
[…]

### 2024-12-03 scripts/02-download_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
-# Purpose: Downloading and saving the data from OpenData Toronto for Police Race and Identity Based Data Collection: Arrests & Strip Searches
+# Purpose: Downloading the data from OpenData Toronto on Police Race and Identity-Based Data Collection: Arrests and Strip Searches.
 # Author: Akshat Aneja
-# Date: 23 September 2024
+# Date: 3 December 2024
 # Contact: akshat.aneja@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: none

### 2024-12-03 scripts/03-clean_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
-# Purpose: Downloading and saving the data from OpenData Toronto for Police Race and Identity Based Data Collection: Arrests & Strip Searches
+# Purpose: Cleaning the data from OpenData Toronto on Police Race and Identity-Based Data Collection: Arrests and Strip Searches.
 # Author: Akshat Aneja
-# Date: 23 September 2024
+# Date: 3 December 2024
 # Contact: akshat.aneja@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: none

### 2024-12-03 scripts/04-test_analysis_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
-# Purpose: Downloading and saving the data from OpenData Toronto for Police Race and Identity Based Data Collection: Arrests & Strip Searches
+# Purpose: Testing the analysis data from OpenData Toronto on Police Race and Identity-Based Data Collection: Arrests and Strip Searches.
 # Author: Akshat Aneja
-# Date: 23 September 2024
+# Date: 3 December 2024
 # Contact: akshat.aneja@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: none
@@ -11,8 +11,6 @@
 #### Workspace setup ####
 library(tidyverse)
 
-
-#### Test data ####
 arrest_data <- read.csv(here::here("data/02-analysis_data/analysis_data.csv"))
 
 # 1. Check if 'arrest_year' is numeric
@@ -42,27 +40,22 @@ valid_arrest_year_values <- all(arrest_data$arrest_year >= 2020 & arrest_data$ar
 # 9. Validate that 'search_reason' values are binary (0 or 1)
 valid_search_reason_values <- all(arrest_data$search_reason >= 0 & arrest_data$search_reason <= 1)
 
-
-
-# 12. Check for missing values in the dataset
+# 10. Check for missing values in the dataset
 no_missing_values <- all(complete.cases(arrest_data))
 
-
-
-
-# 15. Validate 'booked' column as logical or numeric
+# 11. Validate 'booked' column as logical or numeric
 booked_logical <- is.logical(arrest_data$booked) || is.numeric(arrest_data$booked)
 
-# 16. Validate 'search_reason' column as logical or numeric
+# 12. Validate 'search_reason' column as logical or numeric
 search_reason_logical <- is.logical(arrest_data$search_reason) || is.numeric(arrest_data$search_reason)
 
-# 17. Validate that 'arrest_year' does not exceed the current year
+# 13. Validate that 'arrest_year' does not exceed the current year
 valid_years <- all(arrest_data$arrest_year <= as.numeric(format(Sys.Date(), "%Y")))
 
-# 18. Validate the dataset is not empty
+# 14. Validate the dataset is not empty
 non_empty_dataset <- nrow(arrest_data) > 0
 
-# 19. Validate that 'arrest_year' values are above 1900
+# 15. Validate that 'arrest_year' values are above 1900
 valid_arrest_year <- all(arrest_data$arrest_year > 1900)
 
 # Check all conditions

### 2024-12-03 scripts/06-model_data.R

@@ -1,11 +1,12 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Modelling the data from OpenData Toronto on Police Race and Identity-Based Data Collection: Arrests and Strip Searches.
+# Author: Akshat Aneja
+# Date: 3 December 2024
+# Contact: akshat.aneja@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: none
+# Datasets: https://open.toronto.ca/dataset/police-race-and-identity-based-data-collection-arrests-strip-searches/
+
 
 
 #### Workspace setup ####

### 2024-12-03 paper/paper.qmd
Adding pdf of paper and changed beta

--- 2024-12-03 Bias-in-arrests-toronto/paper/paper.qmd
@@
 toc: true
-thanks: "Code and data supporting this analysis is available at: https://github.com/Akshat211202/Analysis-of-bias-in-arrests/"
+thanks: "Code and data supporting this analysis is available at: https://github.com/Akshat211202/bias-in-arrests-toronto"
+
 bibliography: references.bib
@@
 \mbox{logit}(\pi_i) &= \beta_0 + \beta_1 \times \mbox{race}_i + \beta_2 \times \mbox{year}_i \label{eq2} \\
-\beta_0 &\sim \mbox{Normal}(0, 5) \label{eq3} \\
-\beta_1 &\sim \mbox{Normal}(0, 5) \label{eq4} \\
-\beta_2 &\sim \mbox{Normal}(0, 5) \label{eq5}
+\beta_0 &\sim \mbox{Normal}(0, 2.5) \label{eq3} \\
+\beta_1 &\sim \mbox{Normal}(0, 2.5) \label{eq4} \\
+\beta_2 &\sim \mbox{Normal}(0, 2.5) \label{eq5}
 \end{align}
@@
 The Bayesian logistic regression model is represented by Equations (\ref{eq1}), (\ref{eq2}), (\ref{eq3}), (\ref{eq4}), and (\ref{eq5}).
-In these equations: - $y_i$ represents the booking status of the (i)-th individual (0 for not booked and 1 for booked).
-- $\pi_i$ represents the probability of the (i)-th individual being booked after arrest.
-- (\beta\_0) represents the intercept, or the log-odds of being booked for a White individual in the year 2020 (reference group).
-- (\beta\_1) represents how much the log-odds of being booked change on average depending on the race of the individual (relative to a White individual).
-- (\text{race}\_i) represents the race of the (i)-th individual (categorical predictor with levels: White, Black, Indigenous, South Asian, East/Southeast Asian, Middle Eastern, Latino, and Unknown/Legacy).
-- (\beta\_2) represents how much the log-odds of being booked change on average depending on the year (relative to 2020).
-- (\text{year}\_i) represents the year of the (i)-th individual’s arrest (binary: 0 for 2020, 1 for 2021).
+In these equations: $y_i$ represents the booking status of the $i^{th}$ individual (0 for not booked and 1 for booked).
+$\pi_i$ represents the probability of the $i^{th}$ individual being booked after arrest.
+$\beta_0$ represents the intercept, or the log-odds of being booked for a White individual in the year 2020 (reference group).
+$\beta_1$ represents how much the log-odds of being booked change on average depending on the race of the individual (relative to a White individual).
+$\text{race}_i$ represents the race of the $i^{th}$ individual (categorical predictor with levels: White, Black, Indigenous, South Asian, East/Southeast Asian, Middle Eastern, Latino, and Unknown/Legacy).
+$\beta_2$ represents how much the log-odds of being booked change on average depending on the year (relative to 2020).
+$\text{year}_i$ represents the year of the $i^{th}$ individual’s arrest (binary: 0 for 2020, 1 for 2021).
 
-The priors for (\beta\_0), (\beta\_1), and (\beta\_2) are weakly informative normal distributions with a mean of 0 and a standard deviation of 5.
+The priors for $\beta_0$, $\beta_1$, and $\beta_2$ are weakly informative normal distributions with a mean of 0 and a standard deviation of 5.
 These priors allow for flexibility in parameter estimation while regularizing extreme values.
 
-The (\beta) coefficients are important because they provide insight into how an individual’s race and the year of arrest influence the likelihood of being booked after arrest.
+The $\beta$ coefficients are important because they provide insight into how an individual’s race and the year of arrest influence the likelihood of being booked after arrest.
 

### 2024-12-03 paper/paper.qmd
Adding estimand in introduction

@@
 This study seeks to address that gap by analyzing the relationship between ethnicity and booking status.
+
+The study aims to estimate the disparity in arrest rates between racial groups in Toronto, conditioned on demographic and socioeconomic variables such as age, income level, and neighborhood characteristics.
+Specifically, it seeks to quantify the proportion of arrests attributable to racial representation disparities, holding other factors constant.
 

### 2024-12-03 other/datasheet/Datasheet for Police Race and Identity Based Data - Arrests and Strip Searches for the city of Toronto - datasheet_template.qmd
Updating datasheet and references

---
title: "Datasheet for Police Race and Identity Based Data - Arrests and Strip Searches for the city of Toronto"
author: "Akshat Aneja"
thanks: "Code and data are available at: https://github.com/Akshat211202/bias-in-arrests-toronto"
date: today
date-format: long
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1.  *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    -   The dataset was created to investigate arrest patterns in Toronto, focusing on racial representation and potential disparities.
2.  *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    -   The dataset was likely compiled by researchers or academic institutions studying criminal justice or racial equity, potentially in collaboration with public entities.
3.  *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    -   The funding source is not explicitly stated.
4.  *Any other comments?*
    -   None

**Composition**

1.  *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
    -   Each instance represents an individual arrest event, with associated demographic, geographic, and legal attributes.
2.  *How many instances are there in total (of each type, if appropriate)?*
    -   The dataset contains thousands of arrest records (specific count to be derived from the full document).
3.  *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
    -   It is unclear if the dataset includes all arrests or represents a sample.
4.  *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
    -   Each instance includes details such as race, gender, arrest location, type of offense, and socioeconomic indicators of the area.
5.  *Is there a label or target associated with each instance? If so, please provide a description.*
    -   Yes, labels may include arrest outcomes, offense categories, or demographic classifications.
6.  *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
    -   Some records may have missing or incomplete demographic or arrest details.
7.  *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
    -   Relationships such as correlations between demographic variables and arrest outcomes may be implied but not explicitly documented.
8.  *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
    -   No specific splits are provided, but the dataset could be partitioned by time, geography, or demographics.
9.  *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
    -   The dataset may contain biases or inconsistencies due to data collection methods.
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
    -   The dataset is self-contained but may reference external demographic or geographic datasets for enrichment.
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
    -   No personally identifiable information is included; all data appears to be anonymized.
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
    -   No, though the results could highlight sensitive systemic issues.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
    -   Yes, it identifies sub-populations by race, gender, and location.
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
    -   No, all data appears to be anonymized.
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
    -   Yes, data includes variables like race, gender, and arrest location.
16. *Any other comments?*
    -   None

**Collection process**

1.  *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
    -   Data was likely acquired from public law enforcement records and demographic databases.
2.  *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
    -   Data was collected through administrative and observational means, potentially including manual and automated processes.
3.  *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
    -   Unclear; it is not specified if this is a sample or comprehensive dataset.
4.  *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
    -   Likely government or research personnel; details on compensation are unavailable.
5.  *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
    -   Data covers multiple years, though the specific timeframe is not stated.
6.  *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
    -   Not explicitly mentioned.
7.  *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
    -   Data was obtained indirectly from public records and administrative datasets.
8.  *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
    -   No evidence suggests individuals were notified.
9.  *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
    -   No evidence of explicit consent.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
    -   No evidence suggests a consent mechanism was in place.
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
    -   Not mentioned in the document.
12. *Any other comments?*
    -   None

**Preprocessing/cleaning/labeling**

1.  *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
    -   Details about preprocessing or cleaning are not explicitly provided.
2.  *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
    -   Not specified.
3.  *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
    -   Not mentioned.
4.  *Any other comments?*
    -   None.

**Uses**

1.  *Has the dataset been used for any tasks already? If so, please provide a description.*
    -   Yes, for analyzing arrest patterns and racial disparities.
2.  *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
    -   Not specified.
3.  *What (other) tasks could the dataset be used for?*
    -   Policy evaluation, geographic analyses, and socioeconomic impact studies.
4.  *Are there tasks for which the dataset should not be used? If so, please provide a description.*
    -   Should not be used for purposes that perpetuate biases or stigmatize communities.
5.  *Any other comments?*
    -   None.

**Distribution**

1.  *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
    -   Likely available through research publications or institutional platforms.
2.  *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
    -   Via academic platforms or government repositories.
3.  *When will the dataset be distributed?*
    -   Distribution likely coincides with the research publication timeline.
4.  *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
    -   Not explicitly stated, but typically subject to institutional guidelines.
5.  *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
    -   Not mentioned.
6.  *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
    -   Not mentioned.
7.  *Any other comments?*
    -   None.

**Maintenance**

1.  *Who will be supporting/hosting/maintaining the dataset?*
    -   Likely the authors or affiliated academic institutions.
2.  *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
    -   Contact information likely provided in the corresponding research paper.
3.  *Is there an erratum? If so, please provide a link or other access point.*
    -   Not mentioned.
4.  *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
    -   Updates may occur, but no schedule is specified.
5.  *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
    -   Not mentioned.
6.  *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
    -   Not mentioned.
7.  *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
    -   Not mentioned.
8.  *Any other comments?*
    -   None.

\newpage

# References

### 2024-12-03 paper/paper.qmd
Rewrote sections of appendix

--- 2024-12-03 paper/paper.qmd
@@
 ---
-title: "Analyzing Arrest Patterns and Racial Representation in Toronto"
+title: "Assessing Racial Bias in Arrest and Booking Practices in Toronto: Insights and Implications from 2020-2021"
+subtitle: "Black and Indigenous Overrepresentation in Arrests with Stable Booking Probabilities Across Ethnic Groups"
 author: "Akshat Aneja"
-abstract: This paper analyzes arrest data from the Toronto Police Service for the years 2020 and 2021, focusing on the distribution of arrests by ethnic identity, age, and booking status to assess whether systemic bias exists in policing practices.
-The analysis reveals that, while arrests appear to be driven by objective factors, disparities related to socio-economic status may still exist.
-These findings provide important insights that can help improve transparency and accountability in policing, contributing to more equitable law enforcement practices.
+abstract: This study examines arrest and booking data from the Toronto Police Service for 2020 and 2021 to explore potential racial disparities in policing practices.
+While Black and Indigenous individuals were disproportionately represented in arrests, the likelihood of being formally booked after arrest was similar across ethnic groups.
+These findings suggest that disparities may occur earlier in the law enforcement process rather than at the booking stage.
+By highlighting these patterns, the study underscores the need for targeted reforms to ensure equity and transparency in policing practices.
 date: today
@@
 This study seeks to address that gap by analyzing the relationship between ethnicity and booking status.
+
+The study seeks to estimate the relative risk (or odds) of arrest for individuals belonging to specific racial groups compared to a baseline racial group, adjusting for factors such as socioeconomic status, neighborhood characteristics, and crime type.
+This parameter reflects the proportion of observed disparities that can be attributed to racial representation in the city's population.
 
@@
 #| eval: true
-#| fig.cap: "# arrests by Ethnicity in Year 2020 and 2021"
+#| fig.cap: "Number of arrests by Ethnicity in Year 2020 and 2021"
 #| fig.width: 10
@@
   labs(x = "Ethnicity",
-       y = "# Arrests",
+       y = "Number of Arrests",
        fill = "Year") +
@@
 
+\newpage
+
+\appendix
+
 # Appendix
@@
 
-(@fig-ppc) shows posterior predictive check was performed . @fig-ppc illustrates a comparison between posterior distribution simulations and the actual probability of experiencing a custody arrest
+[@fig-ppc] shows a posterior predictive check was performed. It illustrates a comparison between posterior distribution simulations and the actual probability of experiencing a custody arrest
 
@@
 
-In addition, using code adapted from @tellingstorieswithdata, a comparison of the prior and posterior was performed (@fig-comparison).
-@fig-comparison depicts the amounts by which the estimates are modified after considering data [@tellingstorieswithdata].
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-comparison
-#| fig-cap: "Comparison of the prior and posterior"
-
-# The code below was adapted based on Alexander (2023).
-# 
-# posterior_vs_prior(model) +
-#   theme_minimal() +
-#   scale_color_brewer(palette = "Set1") +
-#   theme(legend.position = "none") +
-#   coord_flip()
-```
-
 ### Diagnostics {#sec-diagnostics}
 
-This section assesses whether the `rstanarm` [@citerstanarm] Markov chain Monte Carlo (MCMC) algorithm experienced problems [@tellingstorieswithdata].
-
-Using code adapted from @tellingstorieswithdata, a trace plot was generated (@fig-trace).
-@fig-trace illustrates horizontal lines that are seemingly erratic with chain overlap, indicating nothing unusual (based on @tellingstorieswithdata).
+[@fig-trace] shows well-mixed and overlapping chains, indicating good convergence of the Bayesian logistic regression model's Markov Chain Monte Carlo sampling process.
+This validates the reliability of the posterior estimates reported in the paper, supporting the conclusion that booking probabilities were consistent across ethnic groups and years.
+However, the wide credibility intervals for some coefficients highlight the need for cautious interpretation, especially for smaller ethnic groups, due to limited data and high variability.
 
@@
 
-Moreover, using code adapted from @tellingstorieswithdata, a Rhat plot was created (@fig-rhat).
-@fig-rhat depicts values near 1 and less than or equal to 1.1, indicating no issues (based on @tellingstorieswithdata).
+[@fig-rhat] demonstrates that all Rhat values are approximately 1.0, which indicates that the Markov Chain Monte Carlo (MCMC) chains have converged well.
+This suggests that the Bayesian logistic regression model used in the paper is reliable, as the convergence ensures robust and consistent parameter estimation.
+These results support the validity of the study's findings on arrest and booking probabilities, providing confidence in the statistical inferences drawn from the model.
 
@@
 
-# The code below was adapted from Alexander (2023).
-
 plot(model, "rhat")
@@
 
-Using code adapted from @tellingstorieswithdata, @fig-credibility-intervals-all was generated.
-@fig-credibility-intervals-all illustrates 90\% credibility intervals of all of the predictors of undergoing a custody arrest based on the model from this report.
+The credibility intervals in [@fig-credibility-intervals-all] show that most coefficients, including race and year, have intervals crossing zero, indicating no statistically significant effects of these predictors on booking probabilities.
+This aligns with the paper's findings that booking probabilities were consistent across ethnic groups and years, suggesting limited evidence of systemic bias at the booking stage.
+However, the wide intervals, particularly for smaller ethnic groups like Indigenous and East/Southeast Asian, reflect data limitations and highlight the need for caution in drawing definitive conclusions.
 
@@
 #| label: fig-credibility-intervals-all
-#| fig-cap: "Credibility intervals of all of the custody arrest predictors"
-
-# The code below was adapted from Alexander (2023).
+#| fig-cap: "Credibility Intervals for All Predictors in Booking Analysis"
 
@@
 
-@fig-credibility-intervals-several depicts 90\% credibility intervals of several of the predictors of undergoing a custody arrest.
-In particular, the American Indian/Alaskan Native and Hawaiian/Pacific Islander predictors were omitted for visual clarity (as compared to @fig-credibility-intervals-all).
+The credibility intervals in [@fig-credibility-intervals-several]  focus on a subset of predictors and show that all intervals overlap with zero, indicating no statistically significant effects of these predictors on booking probabilities.
+This reinforces the paper's conclusion that booking probabilities are consistent across ethnic groups and between years, with no strong evidence of systemic bias at the booking stage.
+However, the wide intervals for smaller groups like Indigenous and East/Southeast Asian highlight data limitations and the need for caution in interpreting the results.
 
@@
 #| label: fig-credibility-intervals-several
-#| fig-cap: "Credibility intervals of several of the custody arrest predictors"
-
-# Portions of the code below were adapted from Alexander (2023).
+#| fig-cap: "Credibility Intervals for Key Predictors in Booking Analysis"
 

### 2024-12-04 paper/paper.qmd
Improving grammar

@@
 ---
-title: "Assessing Racial Bias in Arrest and Booking Practices in Toronto: Insights and Implications from 2020-2021"
+title: "Assessing Racial Bias in Arrest and Booking Practices in Toronto from 2020-2021"
 subtitle: "Black and Indigenous Overrepresentation in Arrests with Stable Booking Probabilities Across Ethnic Groups"
 author: "Akshat Aneja"
-abstract: This study examines arrest and booking data from the Toronto Police Service for 2020 and 2021 to explore potential racial disparities in policing practices.
+abstract: This study examines arrest and booking data from the Toronto Police Service for 2020 and 2021 to analyze potential racial disparities in policing practices.
 While Black and Indigenous individuals were disproportionately represented in arrests, the likelihood of being formally booked after arrest was similar across ethnic groups.
@@
 This study examines disparities in booking probabilities across ethnic groups in Toronto during 2020 and 2021.
-While discussions about racial bias in policing often focus on arrest rates or use of force, booking decisions represent another critical stage in law enforcement processes that can reflect systemic bias.
+While discussions about racial bias in policing often focus on arrest rates or use of force, booking decisions represent another important stage in law enforcement processes that can reflect systemic bias.
 
 Booking, the formal process of charging an individual after arrest, determines whether someone enters the criminal justice system.
-Examining booking probabilities provides insight into whether disparities exist in how different racial groups are treated after an arrest.
+Examining booking probabilities provides understanding into whether disparities exist in how different racial groups are treated after an arrest.
 This analysis focuses on identifying patterns in booking decisions for various ethnic groups and whether these patterns changed over the two years.
@@
 Identifying disparities in booking probabilities can help highlight areas where procedural changes may be necessary to ensure fair treatment for all groups.
-This study also explores how booking practices have evolved between 2020 and 2021 to identify potential progress or emerging trends.
+This study also analyzes how booking practices have evolved between 2020 and 2021 to identify potential progress or emerging trends.
 
@@
 
-It is crucial to analyze the distribution of arrests across different age groups to understand potential patterns and disparities.
+It is important to analyze the distribution of arrests across different age groups to understand potential patterns and disparities.
 [@tbl-table1] presents a summary of arrest counts for various ethnic groups across all age categories in 2020 and 2021.
@@
 
-Given the complexity and extensive nature of the dataset, which involves nine age groups across eight ethnicities, summarizing the entire dataset in a comprehensible manner is challenging.
+Given the complexity and extensive nature of the dataset, which involves nine age groups across eight ethnicities, summarizing the entire dataset in a thorough manner is challenging.
 To address this and ensure clarity, we have selected a subset of the data focusing specifically on individuals aged 25 to 34 years.
 This group was chosen as it represents a significant proportion of the data and allows for a more straightforward comparison of arrest trends across ethnic groups while avoiding information overload.
-By narrowing the scope, the analysis becomes more manageable and the insights more focused.
+By narrowing the scope, the analysis becomes more manageable and the findings more focused.
 
@@
 It is important to note that the OpenDataToronto [@rOpenDataToronto] provides population distribution data by ethnicity.
-However, due to the size and diversity of Toronto and its numerous wards, it has been challenging to cross-verify this data with crime rate statistics in a comprehensive manner.
+However, due to the size and diversity of Toronto and its numerous wards, it has been challenging to cross-verify this data with crime rate statistics in a thorough manner.
 Further, a more detailed analysis could be conducted by comparing specific ethnic groups directly.
@@
 
-Referring to [@tbl-table2], [@tbl-table3] and [@fig-fig1] the data reveals a non-uniform distribution of arrests across ethnic groups, with the Black and White populations showing the highest numbers of arrests in both 2020 and 2021.
+Referring to [@tbl-table2], [@tbl-table3] and [@fig-fig1] the data shows a non-uniform distribution of arrests across ethnic groups, with the Black and White populations showing the highest numbers of arrests in both 2020 and 2021.
 Additionally, the data indicates a decline in arrests for these two ethnic groups from 2020 to 2021.
@@
 
-The data reveals a non-uniform distribution of arrests across age groups, with 25 to 34 and 35 to 44 age brackets accounting for the highest number of arrests in both 2020 and 2021.
+The data shows a non-uniform distribution of arrests across age groups, with 25 to 34 and 35 to 44 age brackets accounting for the highest number of arrests in both 2020 and 2021.
 However, there is a notable decrease in the number of arrests for these age groups of 2020 to 2021 which indicates that there may be a possible shift in law enforcement activity.
@@
 
-The $\beta$ coefficients are important because they provide insight into how an individual’s race and the year of arrest influence the likelihood of being booked after arrest.
+The $\beta$ coefficients are important because they provide understanding into how an individual’s race and the year of arrest influence the likelihood of being booked after arrest.
 
@@
 
-The model was designed to provide insights into: Whether racial groups differ in booking probabilities after arrest.
+The model was designed to provide understanding about whether racial groups differ in booking probabilities after arrest.
 Whether booking probabilities changed between 2020 and 2021.
-The Bayesian framework ensures these insights are backed by a robust statistical foundation that accounts for both data-driven and prior uncertainties.
+The Bayesian framework ensures these findings are backed by a robust statistical foundation that accounts for both evidence-based and prior uncertainties.
 
@@
 Wide credibility intervals for racial group coefficients suggest that data limitations (e.g., small sample sizes for some groups) contributed to high variability in results.
-This highlights the need for more comprehensive data to draw robust conclusions.
+This highlights the need for more thorough data to draw robust conclusions.
 
@@
 The predictors in this analysis were ethnicity and year.
-The results provide insights into equity in booking practices and the implications of these findings for law enforcement policies.
+The results provide a good understanding into equity in booking practices and the implications of these findings for law enforcement policies.
 
@@
 
-Analysis of arrest data revealed that Black and Indigenous individuals were overrepresented in arrests relative to their share of the population, consistent with prior findings from the Toronto Police Service and other studies on racial disparities in policing.
+Analysis of arrest data showed that Black and Indigenous individuals were overrepresented in arrests relative to their share of the population, consistent with prior findings from the Toronto Police Service and other studies on racial disparities in policing.
 However, this study focused on booking probabilities after arrest and found no significant differences in booking outcomes across ethnic groups after controlling for year.
@@
 
-Communicating these findings is critical for informed public discourse.
+Communicating these findings is essential for informed public discourse.
 While arrest data often highlights disparities, the results suggest that booking probabilities may not exhibit the same biases.
@@
 The dataset relied on officer-recorded perceived race and binary booking outcomes, omitting variables such as offense type, search status, or socio-economic indicators.
-These omissions may obscure nuanced influences on booking decisions and contribute to the high uncertainty in coefficient estimates for smaller groups.
+These omissions may obscure specific influences on booking decisions and contribute to the high uncertainty in coefficient estimates for smaller groups.
 
@@
 
-1.  **Incorporate Contextual Variables**: Adding variables like offense type or search status would provide a more comprehensive view of factors influencing booking probabilities.
+1.  **Incorporate Contextual Variables**: Adding variables like offense type or search status would provide a more thorough view of factors influencing booking probabilities.
 2.  **Examine Interaction Effects**: Analyzing interactions between race and year could highlight evolving trends for specific groups.
@@
 4.
-**Explore Arrest-to-Booking Pipelines**: Investigating disparities across the entire law enforcement process—from arrests to sentencing—would provide a holistic understanding of equity in policing.
+**Analyze Arrest-to-Booking Pipelines**: Investigating disparities across the entire law enforcement process—from arrests to sentencing—would provide a holistic understanding of equity in policing.
 

### 2024-12-04 paper/paper.qmd
Adding references

@@
 
-## Number of Arrests by Age Group for the year 2020-2021 {#sec-group}
-
-An analysis of arrests by age groups is essential to understanding the demographic patterns of law enforcement activity.
-[@fig-fig3] illustrates the distribution of arrests across different age groups for the years 2020 and 2021.
-The age groups included in the dataset are Under 17, 18 to 24 years, 25 to 34 years, 35 to 44 years, 45 to 54 years, 55 to 64 years, and 65 years and over.
-
-The data shows a non-uniform distribution of arrests across age groups, with 25 to 34 and 35 to 44 age brackets accounting for the highest number of arrests in both 2020 and 2021.
-However, there is a notable decrease in the number of arrests for these age groups of 2020 to 2021 which indicates that there may be a possible shift in law enforcement activity.
-
-In contrast, the other age groups experienced an increase in arrests during the same period.
-In particular, the number of arrests for individuals aged 18 to 24 and 45 to 54 saw an upward trend from 2020 to 2021.
-
 ```{r}
@@
   theme_minimal() +
+  scale_fill_brewer(palette = "Set3") +
   theme(legend.position = "bottom") + 
@@
 ```
+
+## Number of Arrests by Age Group for the year 2020-2021 {#sec-group}
+
+An analysis of arrests by age groups is essential to understanding the demographic patterns of law enforcement activity.
+[@fig-fig3] illustrates the distribution of arrests across different age groups for the years 2020 and 2021.
+The age groups included in the dataset are Under 17, 18 to 24 years, 25 to 34 years, 35 to 44 years, 45 to 54 years, 55 to 64 years, and 65 years and over.
+
+The data shows a non-uniform distribution of arrests across age groups, with 25 to 34 and 35 to 44 age brackets accounting for the highest number of arrests in both 2020 and 2021.
+However, there is a notable decrease in the number of arrests for these age groups of 2020 to 2021 which indicates that there may be a possible shift in law enforcement activity.
+
+In contrast, the other age groups experienced an increase in arrests during the same period.
+In particular, the number of arrests for individuals aged 18 to 24 and 45 to 54 saw an upward trend from 2020 to 2021.
 
@@
 The $\beta$ coefficients are important because they provide understanding into how an individual’s race and the year of arrest influence the likelihood of being booked after arrest.
+
+The `rstanarm` package [@citerstanarm] and the R programming language [@citeR] were employed in order to run this model.
+
+The `modelsummary` package [@citeModelsummary] was used to create [@tbl-model], [@fig-credibility-intervals-all], and [@fig-credibility-intervals-several.]
 
@@
 
-[@fig-ppc] shows a posterior predictive check was performed. It illustrates a comparison between posterior distribution simulations and the actual probability of experiencing a custody arrest
+Using code adapted from [@citeAlexander] a posterior predictive check was performed.
+[@fig-ppc] illustrates a comparison between posterior distribution simulations and the actual probability of experiencing a custody arrest
 
